---
title: "ANSP stats"
output: html_document
date: "2026-05-27"
---
Here is some summary information about the Academy of Natural Sciences at Drexel constituent DB in Neotoma.

```{r setup, include=FALSE}

library(httr)
library(tidyverse)
library(sf)
library(DT)
library(leaflet)
library(tmap)

tables = c("datasets","datasettypes","datasetdatabases","constituentdatabases","sites","collectionunits")

for (q in seq(length(tables))) {
tablename=tables[q]

length_table = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/",tablename,"?count=true")))$data

divider = 80000
num_reps = ceiling(as.numeric(length_table[[1]]$count)/divider)

tablelist = list()
for (i in seq(num_reps)) {
  start= (i-1)*divider
  end = divider
  chunk = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/",tablename,"?count=false&offset=",start,"&limit=",end)))$data
 
  if (!is.null(chunk)) {
  tablelist = append(tablelist,chunk)
  }
  }


table_mat = matrix(nrow=length(tablelist),ncol=length(tablelist[[1]]))

for (i in seq(length(tablelist))) {
  for(j in seq(length(tablelist[[1]])) ) {
    if (!is.null(tablelist[[i]][[j]])) {
      table_mat[i,j] = tablelist[[i]][[j]]
    }}}


table_df = as.data.frame(table_mat)


names(table_df) =names(chunk[[1]])

assign(paste0(tablename,"_df"), table_df)}
```


```{r kdsal, message=FALSE, warning=FALSE,echo=FALSE}

ansp = datasets_df %>% left_join(datasetdatabases_df,by=join_by(datasetid)) %>% left_join(constituentdatabases_df,by=join_by(databaseid)) %>% left_join(datasettypes_df,by=join_by(datasettypeid)) %>% dplyr::filter(databasename == "Academy of Natural Sciences of Drexel University")


ansp_sf = ansp %>% left_join(collectionunits_df,by=join_by(collectionunitid)) %>% left_join(sites_df,by=join_by(siteid)) %>% dplyr::filter(!is.na(longitudewest) & !is.na(latitudesouth)) %>% st_as_sf(coords=c("longitudewest","latitudesouth"),crs=4326)

na_coords = ansp %>% left_join(collectionunits_df,by=join_by(collectionunitid)) %>% left_join(sites_df,by=join_by(siteid)) %>% dplyr::filter(is.na(longitudewest) | is.na(latitudesouth))
```

There is one weird dataset with no coordinates in the associated site, Big Lake:
```{r jsadosi,message=FALSE,warning=FALSE,echo=FALSE}

na_coords %>% select(datasetid,datasetname,notes.x,sitename,latitudenorth,longitudeeast,sitedescription) %>% datatable(rownames=FALSE)
```

Otherwise, here is a count of the datasettypes as well as a table of all the datasets:
```{r jsdfdfdadosi,message=FALSE,warning=FALSE,echo=FALSE}
ansp %>% distinct(datasetid,.keep_all=TRUE) %>% count(datasettype) %>% arrange(desc(n)) %>% datatable(rownames=FALSE)

ansp %>% distinct(datasetid,.keep_all = TRUE) %>% select(datasetid,datasetname,datasettype,description) %>% datatable(rownames=FALSE)

```

And here's an attempt at a map. First one is interactive, includes all sites. Second is static, faceted by datasettype:

```{r klfdsnflksn, message=FALSE,warning=FALSE,echo=FALSE}


pal <- colorFactor(palette = "Set3", domain =ansp_sf$datasettype)


leaflet() %>%
    addTiles() %>%
    addCircleMarkers(data=ansp_sf,
                     stroke=FALSE,
                     color=~pal(datasettype),
                     radius=3,
                     fillOpacity=0.5,
                     popup = ~sitename) %>% 
  addLegend(position = "bottomright",
            pal = pal, 
            values= ansp_sf$datasettype)


library(tmap)

tm_shape(ansp_sf) +
  tm_basemap("OpenStreetMap") +
  tm_dots(
    col = "black",
    size = 0.15,
    alpha = 0.5
  ) +
  tm_facets(by = "datasettype")
```
