---
title: "DPDC Stats"
output: html_document
date: "`r Sys.Date()`"
---

Here is some summary information about the Diatom Paleolimnology Data Cooperative constituent DB in Neotoma.

# Summary of Datasets

```{r setup, include=FALSE}

library(httr)
library(tidyverse)
library(sf)
library(DT)
library(leaflet)
library(tmap)

tables = c("datasets","datasettypes","datasetdatabases","constituentdatabases","sites","collectionunits")

for (q in seq(length(tables))) {
tablename=tables[q]

length_table = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/",tablename,"?count=true")))$data

divider = 80000
num_reps = ceiling(as.numeric(length_table[[1]]$count)/divider)

tablelist = list()
for (i in seq(num_reps)) {
  start= (i-1)*divider
  end = divider
  chunk = content(GET(paste0("https://api.neotomadb.org/v2.0/data/dbtables/",tablename,"?count=false&offset=",start,"&limit=",end)))$data
 
  if (!is.null(chunk)) {
  tablelist = append(tablelist,chunk)
  }
  }


table_mat = matrix(nrow=length(tablelist),ncol=length(tablelist[[1]]))

for (i in seq(length(tablelist))) {
  for(j in seq(length(tablelist[[1]])) ) {
    if (!is.null(tablelist[[i]][[j]])) {
      table_mat[i,j] = tablelist[[i]][[j]]
    }}}


table_df = as.data.frame(table_mat)


names(table_df) =names(chunk[[1]])

assign(paste0(tablename,"_df"), table_df)}
```

```{r kdsal, message=FALSE, warning=FALSE,echo=FALSE}

DPDC = datasets_df %>% left_join(datasetdatabases_df,by=join_by(datasetid)) %>% left_join(constituentdatabases_df,by=join_by(databaseid)) %>% left_join(datasettypes_df,by=join_by(datasettypeid)) %>% dplyr::filter(databasename == "Diatom Paleolimnology Data Cooperative (DPDC)")


DPDC_sf = DPDC %>% left_join(collectionunits_df,by=join_by(collectionunitid)) %>% left_join(sites_df,by=join_by(siteid)) %>% dplyr::filter(!is.na(longitudewest) & !is.na(latitudesouth)) %>% st_as_sf(coords=c("longitudewest","latitudesouth"),crs=4326)

na_coords = DPDC %>% left_join(collectionunits_df,by=join_by(collectionunitid)) %>% left_join(sites_df,by=join_by(siteid)) %>% dplyr::filter(is.na(longitudewest) | is.na(latitudesouth))
```



The below table lists some basic metadata connected to all 680 datasets in the DPDC.

```{r koalnsslkd, message=FALSE,warning=FALSE,echo=FALSE}
DPDC %>% distinct(datasetid,.keep_all = TRUE) %>% select(datasetid,datasetname,datasettype) %>% datatable(rownames=FALSE)

```

The below table represents the same data in a wide format. Instead of each dataset, I list each collectionunit and the number of datasets associated with it by datasettype.

```{r askjldalsk, message=FALSE,warning=FALSE,echo=FALSE}

collectionwide = DPDC %>% mutate(dummyvalue = 1) %>% pivot_wider(id_cols=c(collectionunitid,),names_from=datasettype,values_fn=sum,values_from=dummyvalue)  %>%
  mutate(
    total_datasets = rowSums(across(2:5), na.rm = TRUE)
  ) %>% left_join(collectionunits_df, by=join_by(collectionunitid)) %>% left_join(sites_df,by=join_by(siteid)) %>% select(c(1,7,8,26,2:6,))

datatable(collectionwide,rownames=FALSE)


```

The below table is a count of all the datasettypes associated with the DPDC. The most abundant datasettypes are ```water chemistry``` and ```diatom surface sample```, but there are a wide variety of datasets included which are not as obviously connected to diatoms.

```{r jsdfdfdadosi,message=FALSE,warning=FALSE,echo=FALSE}
DPDC %>% distinct(datasetid,.keep_all=TRUE) %>% count(datasettype) %>% arrange(desc(n)) %>% datatable(rownames=FALSE)
```

The following map shows the spatial distribution of DPDC data by datasettype. You can click a datasettype on or off in order to better see the distribution of the less abundant datasettypes.

```{r klfdsnflksn, message=FALSE,warning=FALSE,echo=FALSE}


pal <- colorFactor(palette = "Set3", domain =DPDC_sf$datasettype)


m <- leaflet() %>%
  addTiles()

for (dt in unique(DPDC_sf$datasettype)) {

  m <- m %>%
    addCircleMarkers(
      data = filter(DPDC_sf, datasettype == dt),
      stroke = FALSE,
      color = pal(dt),
      radius = 3,
      fillOpacity = 0.5,
      popup = ~sitename,
      group = dt
    )
}

m %>%
  addLayersControl(
    overlayGroups = unique(DPDC_sf$datasettype),
    options = layersControlOptions(collapsed = FALSE)
  )

```


# Odd data

There are no datasets associated with sites lacking coordinates.
```{r jsadosi,message=FALSE,warning=FALSE,echo=FALSE}

#na_coords %>% select(datasetid,datasetname,datasettype,sitename,latitudenorth,longitudeeast,sitedescription) %>% datatable(rownames=FALSE)
```

Additionally, I attempted to find any datasets that might be in the DPDC by mistake. I searched specifically for any collectionunits that only had a single dataset, where the dataset was neither explicitly about water chemistry or diatoms. There were no such collectionunits.

```{r psjoaodjps, message=FALSE,warning=FALSE,echo=FALSE}

odd_singletons = collectionwide %>% dplyr::filter(total_datasets == 1 & is.na(`water chemistry`) & is.na(diatom) & is.na(`diatom surface sample`))

datatable(odd_singletons,rownames=FALSE)
```